### 2015-09-21 flux-entropy-paper/flux-entropy-paper.tex
intro updates

\documentclass{sigchi}

%% EXAMPLE BEGIN -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)
% \toappear{Permission to make digital or hard copies of all or part of this work for personal or classroom use is      granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. Copyrights for components of this work owned by others than ACM must be honored. Abstracting with credit is permitted. To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee. Request permissions from permissions@acm.org. \\
% {\emph{CHI'14}}, April 26--May 1, 2014, Toronto, Canada. \\
% Copyright \copyright~2014 ACM ISBN/14/04...\$15.00. \\
% DOI string from ACM form confirmation}
%% EXAMPLE END -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)

% Arabic page numbers for submission.  Remove this line to eliminate
% page numbers for the camera ready copy

\pagenumbering{arabic}

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead 
%\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage[pdftex]{hyperref}
% \usepackage{url}      % llt: nicely formatted URLs
\usepackage{color}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{ccicons}
\usepackage{todonotes}

\usepackage{subcaption}

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={SIGCHI Conference Proceedings Format},
  pdfauthor={LaTeX},
  pdfkeywords={SIGCHI, proceedings, archival format},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
}

% create a shortcut to typeset table headings
% \newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
  \begin{document}

%\title{Quantitative Evaluation of Touch-Screen Group Musical Performance with
 % Transition Matrix Measures}
%\title{Mining interaction log data in long-term practice-led research}

\title{Mining Interaction Log Data in a Creative Arts Practice using Macro Gestures and Transition Matrices}

%in a Creative Arts Practice}

\numberofauthors{3}
\author{%
  \alignauthor{1st Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{City, Country}\\
    \email{e-mail address}}\\
  \alignauthor{2nd Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{City, Country}\\
    \email{e-mail address}}\\
  \alignauthor{3rd Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{City, Country}\\
    \email{e-mail address}}\\
}

\maketitle

\begin{abstract}
  In this note we describe the analysis of a data archive for an
  ensemble musical performance practice on touch screens where
  transition matrix measures are able to distinguish between sessions
  of different contexts, and the internal structure of these sessions.
  The archive includes 95 touch data logs of ensemble touch-screen
  interactions in concerts, rehearslas and formal HCI studies of
  groups of musicians over more than two years.
  The data analysis has been undertaken by first classifying these
  data into higher level gesture types and summarising the performers'
  transitions between these gestures.
  We find that two transition-matrix measure, flux, and entropy,
  can be used to distinguish different performance types and contexts,
  as well as different phases (beginning, middle and end) of all of
  the performances in our corpus.
\end{abstract}

\keywords{  
  Creativity Support Tools;
  Agent; 
  Design;
  Methodology}

\category{H.5.5.}{Information Interfaces and Presentation
  (e.g. HCI)}{Sound and Music Computing}{ Systems}

% \category{See
%   \url{http://acm.org/about/class/1998/} for the full list of ACM
%   classifiers. This section is required.}{}{}

%\section{General comments}
% \begin{itemize}
% \item we should be clear (and careful) where we're talking about
%   groups TMs, and where we're talking about individual performers
% \item what's the terminology we're using for a ``performance''? Is it
%   ``performance'', or is that too music-specific? ``Session''?
%   Alternatives are ``activity'', or ``interaction'', but they're so
%   general\ldots This decision has implications for our other
%   terminology: participant vs performer vs musician, etc.
% \item Title: maybe ``Traces through time: making sense of longitudinal
%   open-ended interaction data'' or some such? I think that's our big
%   contribution: it takes the in-the-wild, practitioner-driven research
%   program stuff and tries to make some quantitative sense out of it
%   post-hoc.
% \end{itemize}

\section{Introduction}

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figures/three-performance-contexts}
  \caption{Collaborative musical interaction with our touch-screen
    interfaces in three different contexts: (left to right) rehearsal
    in a studio,
    performance with percussion instruments, and performance with just
    iPads.\label{fig:three-performance-contexts}}
\end{figure*}

Modern computing devices and interfaces continue to present
opportunities to support new kinds of creative expression and
collaborative interaction~\cite{Resnick:2005yu}. However, it has been
acknowledged that understanding and evaluating interfaces in creative
contexts can be challenging~\cite{Shneiderman:2007qv}. In the context
of live performance, interface evaluation using questionnaires, or
ethnographic data gathering can miss the dynamic need of artists in
the moment of performance and may be impractical in live performative
contexts. When performers use a combuter-based artistic interface,
logged interaction data which has been collected for curation and
preservation~\cite{England:2014ys}, might also allow quantitative
investigation into these performances.

In this note we describe an archive of interaction data logs that has
been accumulated to support a long-term practice-led research program
into simple iPad interfaces for group music-making similar to those
described in Martin et al~\cite{Martin:2014cr}. Over a period of more
than two years (April 2013--June 2015), we collected data from group
music-making sessions where the participants used custom-built,
touch-screen apps on Apple iPad devices. Although several different
apps have been included in our corpus, they shared a common mode of
free-form touch interaction where tapping produced short sounds and
swirling or swiping produced long sounds. Apart from this commonality,
the apps featured a variety of sound palettes, different kinds of
visual feedback, and different kinds of networked interactions
(between the iPads in an ensemble and with a central server).

In all of our interaction sessions, every touch interaction---every
touch down, dragging movement and release---was time-stamped and
logged to disk for later analysis. These sessions included rehearsals
without an audience as well as live performances at arts festivals and
other events. They also included interactions during
formally-structured HCI experiments where participant compared
multiple apps and interfaces. Some of the performances were iPad-only
and others used iPads together with other acoustic percussion
instruments. Although the majority of the sessions were free-form
improvisations, a number of rehearsals and performances involved
composed pieces where the musicians read and performed percussion
notation using the apps.

Representative photographs of some of these session contexts
(rehearsal, performance including other acoustic instruments and iPad
only performance) are are shown
Figure~\ref{fig:three-performance-contexts}) and other descriptive
statistics about the sessions are shown in Table \ref{corpus-table}.
In our corpus, the median length of sessions is 7m29s and the median
number of participants per ensemble is four. Metadata was included
together with the interaction data for each session to record the
\emph{context} (performance, rehearsal, study), session \emph{type}
(improvisation or composition), and \emph{instrumentation}
(iPads-only, or iPads with acoustic percussion). The distribution of
recorded sessions by context and type is shown in Figure
\ref{fig:count-data}.

% It is a feature of creative expression in the arts that highly skilled
% practitioners exploit simple tools to produce works of beauty whether
% it be static paintings or dynamic performance. 
% In previous work~\cite{Martin:2014cr} researchers have shown that
% trained percussionists can exploit a very simple iPad interface to
% discover new categories of musical gestures.
% In that paper, iPad apps
% were studied where tapping was mapped to produce produced short
% sounds, while swiping produced long sounds with the volume
% proportional to the velocity of the moving touch point. 
% It turned out that trained percussionists working together in an
% ensemble were able to discover and repeatedly make use of five macro
% gestures made up of taps and swipes: fast taps, slow taps, random
% taps, short swipes and swirls. This process of discovery of
% interesting musical gestures from a simple interface is analogous to
% what percussionists do with acoustic instruments -- drums and bowls
% and bells are often tapped, stroked and hit in novel ways to
% communicate meaning and emotion through music~\cite{Schick:2006fk}.

% TODO fix this table to remove descriptive data about the
% demonstration sessions.
\begin{table}
\centering
\begin{tabular}{l|llll}
\hline
            & Total & Min  & Median   & Max     \\ 
\hline
N           & 150       &         &        &  \\
%Record size & 306MB     &         &        &   \\
Length      & 20H34M22S & 1M0S    & 7M29S  & 26M0S\\
Participants& 524       & 1       & 4      & 9    \\
Flux        &           & 0.02239 & 0.1445 & 0.3246\\
Entropy     &           &  &  & \\          
\hline
\end{tabular}
\caption{
  Representative data about the corpus of musical touch-screen
  interaction sessions used in this study. These sessions were
  rehearsals, performances, or studies and 
  participants improvised or performed compositions. Each record
  consisted of touch-data recorded in CSV format and was later analysed
  with a gestural classification system. Some participants were
  present in multiple sessions, however this data was not tracked in
  the corpus.\label{corpus-table}}
\end{table}

\section{Macro Gestures, Transition Matrices and Matrix Measures}

A useful way of modelling user interfaces is to consider their
configurations as state vectors and to deploy the methods of matrix
algebra to describe transitions between these states during an
interaction sequence~\cite{Thimbleby:2001kq, Thimbleby:2004fj}.
An opposite approach is to record users' interactions with an
interface, manually coding interaction states. 
This method has been applied to usability analysis of resource
planning software for security applications~\cite{Kannampallil:2007fp}
where the transition probability matrix of these state sequences
was used to draw conclusions about the nature of this interactions.
In the creative arts, editing commands at a computer interface for the
``live coding'' of computer music has also been coded as states and
analysed using transition matrices~\cite{Swift:2014tya}, in both the
textual and musical dimensions, in order to arrive at conclusions
regarding artistic style.

In our corpus of iPad musical performances, rehearsals and studies, we
have used the continuous touch-screen gestures identified by Martin et
al~\cite{Martin:2014cr} as a basis for a musical state-space.
Each performer in each musical piece can then be thought of as
producing a ``score'' of state transitions and an overall transition
matrix for the entire ensemble can be derived by averaging the
individual transition matrices.

Suppose that each participant's activity in a collaborative
interaction can at any given time be classified as belonging to one of
$m$ distinct \emph{states}. Over the course of the activity, each
participant $X$ generates a length $N$ sequence of these interaction
states
\begin{equation}
 X_n \hskip 2em n = 1, \ldots, N
\end{equation}
which can be thought of as Markov process. If there is an (automated)
classification process for extracting each participant's state
sequence from (or during) their interaction data, then we can estimate
the $m \times m$ transition matrix $P$ which characterises this
process by simply counting the number of times state $j$ follows state
$i$. The entries of $P$ are then given by:
$p_{ij} = \frac{N_{ij}}{\sum_i \sum_j N_{ij}}$.
% \begin{equation}
%   p_{ij} = \frac{N_{ij}}{\sum_i \sum_j N_{ij}}
% \end{equation}
The transition activity of the whole ensemble can be calculated by
averaging the TM for each performer. The matrix can be normalised so
that the sum of all elements is equal to one.

Previous work~\cite{Martin:2015jk} showed that the raw touch data
obtained from interaction with simple iPad interfaces similar to ours
could be segmented using a Random Forest\cite{Breiman:2001kx}
classifier\footnote{The classifier used was from Python's Scikit-learn
  package\cite{scikit-learn}.}. The classifier was applied to 5-second
windows of the touch-data to provide a chain of gesture-states for
each participant in each session. These chains can be used to create
$9 \times 9$ (the classifier recognises 9~distinct gestures)
transition matrices for individual performers and the whole ensemble
in each session.

\section{Quantitative transition matrix statistics}
\label{sec:underst-impr-group}

While the interaction sessions recorded in our corpus were performed
by many different participants with varying experience levels, in
different contexts and on different instrumental setups, the common
interface and method of recording the data allows all the sessions to
be transformed into gesture sequences. These transition matrices have
been used as a visualisation technique to reveal qualitative
differences in structure in improvisational
music-making\cite{Swift:2014tya}. In this paper, however, we wish to
extract quantitative statistics from these TMs for comparing and
tracking different sessions over time.

To do this we calculate two high-level quantities from each session:
``flux'' and ``entropy''. The flux of a sequence is defined as the
ratio of state transitions (e.g. $A \rightarrow B$ where $A \neq B$),
to self transitions (e.g. $A \rightarrow A $). The flux of a TM $P$ is
given by: $\mathrm{flux}(P) = 1 - \mathrm{tr}(P)$. Where the trace
$\mathrm{tr}(P)$ is the sum of the diagonal entries of $P$.
Intuitively, flux is a measure of how frequently the participant/group
changes state. Flux returns a value in the interval $[0,1]$ and will
return 0 when participants never change state, and 1 when participants
never stay in the same state for two measurements in a row.

Another useful measure that can be used on a transition matrix is its
entropy, defined in the information-theoretic\cite{Shannon:1948rt}
sense: $H(P) = -\sum_{i,j}p_{ij}\log_2(p_{ij})$.
% \begin{equation}
%   H(P) = -\sum_{i,j}p_{ij}\log_2(p_{ij})
% \end{equation}
This entropy measure is small when the matrix is sparse, and largest
when each matrix element is equal. It offers a different perspective
on collaborative interactions than the flux measure by capturing the
breadth of the gestural space explored by the participants throughout
the course of an interaction. Consider the degenerate case of a
participant who alternates between two states for a whole performance:
the flux in this case will be maximal ($\mathrm{flux}(P) = 1$) since
there are no self-transitions, (only $A \rightarrow B$ or
$ B \rightarrow A$) even though the participant has only used a small
subset of the state space. The entropy of this interaction, on the
other hand, will be low. Entropy, therefore, is a measure of how broad
the participant's exploration of the state space is in a given
interaction.

In the rest of this paper we will use these matrix measures to answer
two questions. First, do the transition matrix measures allow us to
distinguish between different performance contexts, types and
instrumental setups? Secondly, do these measures allow us further
insight into the internal structure of performances?

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/sessions-count}
  \caption{Distribution of the sessions in our interaction corpus by
    performance context. Improvisations or compositions were repeated
    several times in rehearsals and study sessions but only performed
    once in concerts.
    \label{fig:count-data}}
\end{figure}

\subsection{Differentiating interaction sessions}
\label{differentiating-interaction-sessions}

\begin{figure} \centering
  \includegraphics[width=\linewidth]{figures/context-flux-entropy-boxplot}
  \caption{Distributions of flux and entropy values by performance
context. Study and rehearsal sessions had the highest flux for
improvisations. Compositions had lower
flux in rehearsals than performances.
\label{fig:flux-entropy-boxplot}}
\end{figure}

The distribution of flux and entropy with respect to session contexts,
types and instrumentation are shown in Figure
\ref{fig:flux-entropy-boxplot}. A three-way ANOVA procedure was
performed on the data set to find whether the metatdata could
significantly predict the entropy and flux values.

All main effects on flux were found to be significant: performance
context ($F(2,86) = 6.28, p < 0.01$), performance type
($F(1,86) = 7.21, p < 0.01$), and instrumentation
($F(1,86) = 20.06, p < 0.001$). Significant interaction effects were
also found for performance context and type
($F(1,86) = 6.18, p < 0.05$), and performance context and
instrumentation ($F(1,86) = 10.48, p < 0.01$).

For entropy, significant main effects were found due to performance
context ($F(2,86) = 12.10, p < 0.001$), and instrumentation
($F(1,86) = 25.86$). The interaction of performance type and
instrumentation was also found to be significant
($F(1,86) = 9.93, p<0.01$).

These tests show that the different interaction styles that might be
present in performance contexts, types and instrumentations are
reflected in the values of flux and entropy of their transition
matrices.

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.5\linewidth}
        \centering
          \includegraphics[width=\linewidth]{figures/type-section-flux-entropy}
          \caption{Performances divided by performance type. \label{fig:type-section-flux-entropy}}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\linewidth}
        \centering
          \includegraphics[width=\linewidth]{figures/context-section-flux-entropy}
          \caption{Sessions divided by performance context. \label{fig:context-section-flux-entropy}}
    \end{subfigure}
    \caption{The flux and entropy of different interaction sessions
      divided into beginning, middle, and end.}
\end{figure*}

\subsection{Breaking performances into sections}

It is clear from these results that flux and entropy can be used to
differentiate between performance styles and types, but can they be
used to understand some of the internal structure of performances? A
simple model for understanding temporal art-forms is the ``beginning,
middle, end'' structure. To investigate how this structure fits our
corpus, we can divide the gestural-sequences of each session into
thirds and calculate the transition matrices given by each section.

Figure~\ref{fig:type-section-flux-entropy} shows that the flux and
entropy of compositions are much more stable with respect to section
than for improvisations. The stability of compositions is to be
expected since performers are expected to use the same gestures in
each rehearsal and performance of a composition, it may be that
different kinds of compositions could result in different structural
differences. Both measures have a downward trend through sections in
improvisations with Kruskal-Wallis tests showing a significant effect
of section on flux ($\chi^2(2)=5.8, p=0.05$). This could be explained
by performers entering an exploration stage at the start of an
improvisation, where they change gesture frequently to experiment with
different ideas. At the end of an improvisation, performers may change
gesture less frequently while ``winding up'' the performance.

The gestural change in improvisation can be further explored by
dividing the sessions by performance context. Figure
\ref{fig:context-section-flux-entropy} shows this division for
improvisations only. In rehearsals and studies the distribution of
flux and entropy over sections has a similar shape, both measures drop
in the ending section of studies and after the beginning section in
rehearsals. In performances however, flux drops after the beginning
while entropy drops substantially in the ending section. This
indicates a more constrained performance style in this closing
section, where performers transition between fewer gestures in the
final section of performances.

% todo add the plots here.

% \begin{figure}
%   \centering
%   \includegraphics[width=\linewidth]{figures/type-section-flux}
%   \includegraphics[width=\linewidth]{figures/type-section-entropy}
%   \caption{The flux of compositions and improvisations divided into
%     beginning, middle, and end. Composition flux remains stable throughout,
%     but improvisations show a downward trend and composition entropy tends to move up
%     at the end, while flux moves down.
%     \label{fig:type-section-flux-entropy}}
% \end{figure}

% \begin{figure}
%   \centering
%   \includegraphics[width=\linewidth]{figures/context-section-flux}
%   \includegraphics[width=\linewidth]{figures/context-section-entropy}
%   \caption{The distribution of flux and entropy in performances divided by ``beginnings,
%     middles, and ends''.
%     \label{fig:context-section-flux-entropy}}
% \end{figure}

\subsection{Performance styles over time}

While the interaction session in our corpus have many different
contexts, they are all part of a program of artistic and HCI research
that has evolved over time. Figure \ref{fig:flux-entropy-through-time}
shows the values of flux and entropy for each session in the corpus.
Over the course of this research program, different ensembles and
different activities have been emphasised, some of these differences
can be seen as clusters of performances with similar flux and entropy
in the two plots. For instance, the earliest series of rehearsals and
performances had flux between 0.05 and 0.15, while the period of
activity after July 2014 had flux between 0.1 and 0.25. This change
corresponded with an intensive series of study sessions and rehearsals
where the participants developed a more ``fluctuating'' style of
performance. The entropy between these two periods, however, remains
at a similar level, showing the the space of available gestures that
the performers explored was relatively stable.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/flux-entropy-through-time}
  \caption{Distribution of flux and entropy values of sessions through time.
    Different performance styles are visible as clusters of similar
    flux regardless of performance context.
    \label{fig:flux-entropy-through-time}}
\end{figure}

\section{Conclusions}

In this paper we have extended our previous work on transition
matrices with two simple matrix measures, flux and entropy, both of
which have been demonstrated to have significant discriminatory power
with regards to a corpus of 150 collaborative touch-screen
interactions. Not only do these measures differentiate between
different types of sessions and the internal structure of these
sessions, but they have explanatory power that has thrown new light on
the musical interactions under examination. HCI researchers are
increasingly concerned with interactions that happen in-the-wild,
where the interactions of many users may be collected without the
opportunity to survey or interview users. We suggest that analysing
interaction-state transitions using matrices and simple measures can
lead to new insights into HCI interactions outside the lab.

%TODO Mention Manovich transmutation stuff
The macro-gestures that these processes reveal could also be used as
the basis for new works of new media art~\cite{Manovich:2002ly}.

% Additionally, a wider dissemination of such log data will better
% support the replication of findings in HCI
% research~\cite{Wilson:2011ve}

%\pagebreak
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{references}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: