### 2016-01-04 CHI2016-DataMiningPaper-Rebuttal.md
updates with CHI2016 rebuttal

Thanks to the reviewers and ACs for their careful consideration, in particular, we wish to thank R2 for extremely insightful and useful comments. We think that we can address R2's criticisms to allow our findings to be more confidently interpreted. We will also address suggested changes from all of the reviewers to improve clarity and add a citation of Ge Wang's related work.

In our note, we showed that interaction-protocol collection and analysis can yield interesting insights into artistic processes where other evaluations would be impractical. Studying a dataset from a real performance practice is called "significant and realistic" by R3 who noted that our findings are a "concrete research contribution". Other reviewers were also positive and we hope that given our response to the specific criticisms our note will be accepted.

It should be mentioned that most objections refer to compositions, but these data are less important than the improvisations in our analysis and in the note's contribution. We acknowledge that the research questions could be more prominently situated and we will move them from page 3 to the front matter of the note.

# Reviewers' major questions:

- Effect of Tempo (R2):

R2 is correct in suggesting that a kind of "gestural tempo" is encoded in our transition matrices by the timing of the gestural classifier (1s steps, 5s window), however we argue that this is not a significant problem for interpreting the data.

Most of our data relates to improvisations where tempo would affect the internal speed of gestures but not the rate of transition. In this case rate of transition through different gestures is equivalent to performers' "adventurousness" in the gestural space. The compositions used in our research were certainly played in different tempi which may have affected the gestural tempo, but our analysis suggests that the session-lengths of compositions did not significantly correlate with flux or entropy.

We will include these points in our discussion but our claim - that flux and entropy can differentiate between performance types and contexts – still holds.

- Distributions of Compositions (R2):

Three different compositions were present in the data. Out of 23 composition sessions, 17 were of one piece, with the other two accounting for 4 and 2 sessions. For the first two compositions, the distribution was close to 25% concerts, 75% rehearsals, and the last had one rehearsal and one performance. We will clarify that the distribution between concert and rehearsal sessions was quite similar across all three compositions, so that the change in session type is the most likely source of the change in Flux and Entropy between the performance contexts.

- Transition Matrices over Ensembles/Individuals (R2):

We agree that averaging matrices for ensembles results in an overview statistic. In Figure 3 and Figure 5, the statistics have been calculated for ensembles as this is more appropriate for the broader differentiation between performance contexts and types. In Figure 4, they are calculated for individual performers to analyse local behaviour inside each session. We will state this difference clearly in a revised note.

- Goal of the "Breaking Sessions into Sections" analysis (R2):

This section and Figure 4 are mainly focussed on improvisations. The goal is to investigate the internal structure of improvised playing and to see if these sessions do show evidence of a beginning/middle/end structure which we have shown statistically for flux. Compositions are included in Figure 4a to contrast with the clear structure of the improvisations. Even though, as R2 has suggested, the thirds division may not be optimal for identifying segments, structural effects can still be seen in the figures and in the results of a Kruskal-Wallis test for flux. Knowing more about the structure of improvisations will be useful for educational work with these apps, and for future app designs which might e.g. attempt to respond in particular ways in an "ending" phase. Future work could determine optimal divisions inside session.

- Why these Metrics? (R1,R3):

R1 has suggested some alternative metrics for measuring transitions and asks for further discussion, while R3 asks for more theoretical motivation. The advantage of flux and entropy is that they are simple to calculate and can be motivated with the intuitive performative examples described in the note. Simonton's "melodic originality" metric is 
defined by a "ground truth" corpus of 15000 melodies by 500 notable classical composers. It would be difficult to motivate such a definition from our more modest corpus. Kerne et al.'s "ideation metrics" require a classification of an "idea" in the data - single transitions in our data are too low level to be mapped into this system. Our metrics allow very low-level data to be investigated with minimal interpretation. We will cite these very interesting references to help situate our note.